DANA Normalization Assessment for the Paired MSK Data Sets

### 2021-12-17 MSK_Data_Analysis.Rmd
Major Revision
Considerably improved the DANA method by replacing the metric for removal of handling effects by the mean-squared correlation reduction in negative controls. Updated figures and results

---
title: "Normalization Performance Analysis for the Paired MSK Data Sets"
author: "Yannick Dueren, Johannes Lederer, Li-Xuan Qin"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
# abstract: "TODO: This is the abstract."
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "results") })
---

@@

The settings list `settings` configure which parts, simulations, and analysis steps of this R markdown are executed. This way single components of the analysis can be disabled for the purpose of saving computation time. Setting `settings$all` to `TRUE` will set all other options to true regardless of their value.

```{r analysis.settings}
settings <- list(
  # If true sets all other settings to TRUE
  all=F,
  # Application of the full DANA pipeline to the benchmark data
  bench.DANA=T,
  # Application of the full DANA pipeline to the test data
  test.DANA=T,
  # Generate interactive plots of co-expression structures
  investigate.coexpression=T,
  # Comparison of coexpression estimation methods
  compare.corr.methods=T,
  # Full Pipeline comparison of two correlation methods
  compare.two.methods=F,
  # Compare DANA results for different cutoff choices
  compare.cutoffs=T,
  # Perform Differential Expression Analysis for test and benchmark data sets
  perform.DEA=T,
  # Generate paper figures
  generate.paper.figs=T,
  # Specify if paper figures are exported
  export.figures = F,
  # Path for file exports
  paper.fig.path = "../danawriteup/figs/",
  # Clears environment
  debug=TRUE
)
if(settings$all) {
  settings$bench.DANA=TRUE
  settings$test.DANA=TRUE
  settings$investigate.coexpression=TRUE
  settings$compare.corr.methods=TRUE
  settings$compare.two.methods=TRUE
  settings$generate.paper.figs=TRUE
}
if(settings$generate.paper.figs) {
  settings$bench.DANA=TRUE
  settings$test.DANA=TRUE
}
if(settings$debug) rm(list=ls()[ls()!="settings"])
```

@@

  # # Display mean correlation reduction in negative controls
  # compare.test.mcr <- cbind(DANA.test.mb.bic.res$metrics$mcr,
  #                           DANA.test.mb.aic.res$metrics$mcr,
  #                           DANA.test.mb.cv.res$metrics$mcr,
  #                           DANA.test.mb.av.res$metrics$mcr,
  #                           DANA.test.glasso.ric.res$metrics$mcr,
  #                           DANA.test.glasso.stars.res$metrics$mcr,
  #                           DANA.test.huge.mb.ric.res$metrics$mcr,
  #                           DANA.test.huge.mb.stars.res$metrics$mcr,
  #                           DANA.test.fastggm.res$metrics$mcr,
  #                           DANA.test.pearson.res$metrics$mcr,
  #                           DANA.test.spearman.res$metrics$mcr)
  # rownames(compare.test.mcr) <- rownames(DANA.test.mb.bic.res$metrics)
  # colnames(compare.test.mcr) <- c("mb.bic", "mb.aic", "mb.cv", "mb.av", "glasso.ric", "glasso.stars", "huge.mb.ric", "huge.mb.stars", "fastggm", "pearson", "spearman")
  # stargazer(compare.test.mcr, type="text", summary=FALSE, digits=5, 
  #             title="Test data - mcr-", align=TRUE)
  # corrplot.mixed(cor(compare.test.mcr),lower="circle",upper="number")

@@

In this section, we compare two methods for estimating the co-expression of miRNAs to each other. The methods can be chosen individually to positive and negative controls and include all available marginal and partial correaltion methods.

@@

```{r DANA.test.different.cutoffs}
if(settings$compare.cutoffs) {
  ## Configuration of three possible cutoff choices
  cutoff.conf1 <- list(
    t.zero = 2,  # zero-expressed in [0, t.zero)
    t.poor = 10,  # poorly-expressed in [t.zero, t.poor]
    t.well = 2^7  # well-expressed in [t.well, inf)
  )
  cutoff.conf2 <- list(
    t.zero = 1.5,  # zero-expressed in [0, t.zero)
    t.poor = 5,  # poorly-expressed in [t.zero, t.poor]
    t.well = 2^6  # well-expressed in [t.well, inf)
  )
  cutoff.conf3 <- list(
    t.zero = 3,  # zero-expressed in [0, t.zero)
    t.poor = 16,  # poorly-expressed in [t.zero, t.poor]
    t.well = 2^8  # well-expressed in [t.well, inf)
  )

@@

  # Combined read count histogram for test and benchmark data
  df <- data.frame(bench=bench.RC.log2, test=test.RC.log2)
  p.RC.hist <- ggplot(df) + 
    theme_classic() +
    geom_histogram(aes(x=test, y=..count..,fill="#69b3a2"), binwidth = .1) +
    geom_histogram(aes(x=bench, y=-..count.., fill="#404080"), binwidth = .1) +
    geom_vline(xintercept = log2(config$t.zero+1), color="#5851b8", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.poor+1), color="#5851b8", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.well+1), color="#E7298A",  linetype="longdash") +
    geom_segment(aes(x = log2(config$t.zero+1), y = 75, xend = log2(config$t.poor+1), yend = 75),
                 arrow=arrow(length=unit(.07, "in"), ends="both"),
                 color="#5851b8") +
    annotate(geom="label", x=(log2(config$t.zero+1)+log2(config$t.poor+1))/2.0, y=84,
             label="neg. contr.", color="#5851b8", size=4, fill = alpha(c("white"),0.85), label.size = NA) +
    geom_segment(aes(x = log2(config$t.well+1), y = 75, xend = 10, yend = 75),
                 arrow=arrow(length=unit(.07, "in"), ends="last"),
                 color="#E7298A") +
    annotate(geom="label", x=log2(config$t.well+1)+.5, y=84,
             label="pos. contr.", color="#E7298A", size=4, hjust=0, fill = alpha(c("white"),0.85), label.size = NA) +
    # ylim(c(-90,90)) +
    scale_fill_manual(name="Data set",values=c("#404080", "#69b3a2"), labels=c("benchmark", "test"), guide = guide_legend(reverse=TRUE)) +
    # ggtitle("log2(Read count + 1)") +
    theme(legend.position=c(0.86,0.86), 
          # legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.9))) +
    scale_y_continuous(breaks=c(-50,0,50), labels=c(50,0,50), limits=c(-95,95)) +
    xlab("Mean log2(read count)") +
    ylab("Frequency")
    # theme(panel.border = element_blank(),
    #       axis.line = element_line(color = 'black'))
  print(p.RC.hist)
  rm(df)

@@

  # Common control miRNAs
  pos.controls.bench <- rownames(bench.model$pos$corr)
  pos.controls.test <- rownames(test.model$pos$corr)
  pos.controls.common <- intersect(pos.controls.test, pos.controls.bench)
  neg.controls.bench <- rownames(bench.model$neg$corr)
  neg.controls.test <- rownames(test.model$neg$corr)
  # neg.controls.common <- intersect(neg.controls.test, neg.controls.bench)
  # reduce the number of genes for the plot
  pos.controls.plot <- pos.controls.common[2:num.genes.plot.pos+1]

@@

  ### Generate plots
  # Positive controls
  p.corr.pos.bench <- ggplot.corr(corr.bench.pos, 
                                  clusters=clusters, 
                                  threshold=0, 
                                  title="benchmark (un-normalized)")
  p.corr.pos.test <- ggplot.corr(corr.test.pos, 
                                 clusters=clusters, 
                                 threshold=0, 
                                 title="test (un-normalized)")
  p.corr.pos.test.norm <- ggplot.corr(corr.test.norm.pos, 
                                     clusters=clusters, 
                                     threshold=0, 
                                     title="test (TMM)")

  p.corr.pos <- arrangeGrob(p.corr.pos.bench + theme(legend.position="none"), 
                            p.corr.pos.test + theme(legend.position="none"),
                            get.legend(p.corr.pos.bench), 
                            widths = c(2,2,1))
  grid.arrange(p.corr.pos)  # plot

  p.corr.pos.three <- arrangeGrob(p.corr.pos.bench + theme(legend.position="none"), 
                                  p.corr.pos.test.norm + theme(legend.position="none"),
                                  p.corr.pos.test + theme(legend.position="none"),
                                  get.legend(p.corr.pos.bench + theme(legend.position = "bottom")),
                                  layout_matrix=rbind(c(1,2,3), c(4,4,4)),
                                  heights = c(5,1))
  grid.arrange(p.corr.pos.three)  # plot

@@

  p.corr.neg <- arrangeGrob(p.corr.neg.bench + theme(legend.position="none"), 
                            p.corr.neg.test + theme(legend.position="none"),
                            get.legend(p.corr.neg.bench),
                            widths = c(2,2,1))
  grid.arrange(p.corr.neg)  # plot

  p.corr.neg.three <- arrangeGrob(p.corr.neg.bench + theme(legend.position="none"), 
                                  p.corr.neg.test.norm + theme(legend.position="none"),
                                  p.corr.neg.test + theme(legend.position="none"),
                                  get.legend(p.corr.neg.bench + theme(legend.position = "bottom")),
                                  layout_matrix=rbind(c(1,1,2,2,3,3), c(4,4,4,4,4,4)),
                                  heights = c(5,1))
  grid.arrange(p.corr.neg.three)  # plot

@@

```{r generate.paper.figs.corr.density.paper}
if(settings$generate.paper.figs) {
  # Co-expression models
  bench.model <- DANA.bench$data.model
  bench.norm.model <- DANA.bench$norm.models$benchmark.TMM
  test.model <- DANA.test$data.model
  test.norm.model1 <- DANA.test$norm.models$test.TMM
  test.norm.model2 <- DANA.test$norm.models$test.RUVs

  # Set number of genes for negative correlation to minimum
  num.genes.plot.neg <- min(dim(test.model$neg$corr)[1],
                            dim(bench.model$neg$corr)[1],
                            dim(test.norm.model1$neg$corr)[1],
                            dim(test.norm.model2$neg$corr)[1])

  # Subsetted correlation matrices
  corr.bench.neg <- bench.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.bench.norm1.neg <- bench.norm.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.neg <- test.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm1.neg <- test.norm.model1$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm2.neg <- test.norm.model2$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]

  ## extract non-zero correlations
  nnz.corr.test.neg <- corr.test.neg[upper.tri(corr.test.neg)]
  nnz.corr.test.neg <- nnz.corr.test.neg[nnz.corr.test.neg != 0]
  nnz.corr.test.norm1.neg <- corr.test.norm1.neg[upper.tri(corr.test.norm1.neg)]
  nnz.corr.test.norm1.neg <- nnz.corr.test.norm1.neg[nnz.corr.test.norm1.neg != 0]
  nnz.corr.test.norm2.neg <- corr.test.norm2.neg[upper.tri(corr.test.norm2.neg)]
  nnz.corr.test.norm2.neg <- nnz.corr.test.norm2.neg[nnz.corr.test.norm2.neg != 0]
  nnz.corr.bench.neg <- corr.bench.neg[upper.tri(corr.bench.neg)]
  nnz.corr.bench.neg <- nnz.corr.bench.neg[nnz.corr.bench.neg != 0]
  nnz.corr.bench.norm1.neg <- corr.bench.norm1.neg[upper.tri(corr.bench.norm1.neg)]
  nnz.corr.bench.norm1.neg <- nnz.corr.bench.norm1.neg[nnz.corr.bench.norm1.neg != 0]

  # Keep positive correlations
  nnz.corr.test.neg <- nnz.corr.test.neg[nnz.corr.test.neg > 0]
  nnz.corr.test.norm1.neg <- nnz.corr.test.norm1.neg[nnz.corr.test.norm1.neg > 0]
  nnz.corr.test.norm2.neg <- nnz.corr.test.norm2.neg[nnz.corr.test.norm2.neg > 0]
  nnz.corr.bench.neg <- nnz.corr.bench.neg[nnz.corr.bench.neg > 0]
  nnz.corr.bench.norm1.neg <- nnz.corr.bench.norm1.neg[nnz.corr.bench.norm1.neg > 0]

  ## direct comparison of negative controls benchmark vs test vs test.norm1
  neg.corr <- data.frame(
    control=factor(c(rep("Benchmark", length(nnz.corr.bench.neg)), 
                     rep("Test", length(nnz.corr.test.neg)),
                     rep("Test (TMM)", length(nnz.corr.test.norm1.neg)))),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.test.neg, nnz.corr.test.norm1.neg))
    )

  p.density.neg.nnz.freq <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=.9) +
    theme_classic() +
    xlim(0, 1) +
    scale_linetype_manual(values=c("solid", "solid", "twodash"))+
    scale_color_manual(values=c('#1B9E77','#D95F02','#D95F02')) + # First 2 colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.85,0.9), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation") +
    ggtitle("Positive correlation among negative controls")
  print(p.density.neg.nnz.freq)

  ## direct comparison of negative controls benchmark vs test vs test.norm1 vs test.norm2
  neg.corr <- data.frame(
    control=factor(c(rep("Benchmark", length(nnz.corr.bench.neg)), 
                     rep("Test", length(nnz.corr.test.neg)),
                     rep("Test (TMM)", length(nnz.corr.test.norm1.neg)),
                     rep("Test (RUVs)", length(nnz.corr.test.norm2.neg)))),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.test.neg, nnz.corr.test.norm1.neg, nnz.corr.test.norm2.neg))
    )
  # ttt <- data.frame(prop.table(table(method=neg.corr$control, corr=neg.corr$corr), 1))

  p.density.neg.nnz.freq <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=.9) +
    theme_classic() +
    xlim(0, 1) +
    # scale_linetype_manual(values=c("twodash", "solid", "solid", "solid"))+
    scale_color_manual(values=c('#e7298a','#1B9E77','#D95F02', '#7570B3')) + # First colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.85,0.86), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    theme(legend.key.width = unit(1.25,"cm")) +
    ylab("Frequency") +
    xlab("Marginal correlation") #+ ggtitle("Positive correlation among negative controls")
  print(p.density.neg.nnz.freq)

  # Additional Plot with normalized benchmark
  neg.corr <- data.frame(
    control=factor(c(rep("Benchmark", length(nnz.corr.bench.neg)), 
                     rep("Benchmark (TMM)", length(nnz.corr.bench.norm1.neg)),
                     rep("Test", length(nnz.corr.test.neg)),
                     rep("Test (TMM)", length(nnz.corr.test.norm1.neg)))),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.bench.norm1.neg, nnz.corr.test.neg, nnz.corr.test.norm1.neg))
    )
  p.density.neg.nnz.freq.four <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=1) +
    theme_classic() +
    xlim(0, 1) +
    theme(legend.position=c(0.9,0.9), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation") +
    scale_linetype_manual(values=c("solid", "twodash","solid", "twodash"))+
    scale_color_manual(values=c('#1B9E77','#1B9E77','#D95F02','#D95F02'))  + # First 2 colors from rcolorbrewer set "Dark2" 
    ggtitle("Positive correlation among negative controls")
  print(p.density.neg.nnz.freq.four)

  if(settings$export.figures) {
    # ggsave(paste0(settings$paper.fig.path, "MSK_CorrDensityNeg_Hist.pdf"), p.density.neg.nnz.hist, width = 5, height=4, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "MSK_CorrDensityNeg_Freq.pdf"), p.density.neg.nnz.freq, width = 5, height=4, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "MSK_CorrDensityNeg_Freq_WithBench.pdf"), p.density.neg.nnz.freq.four, width = 5, height=4, device="pdf")
  }
}
```

@@

```{r generate.paper.figs.corr.density.supplements}
if(settings$generate.paper.figs) {
  # Co-expression models
  bench.model <- DANA.bench$data.model
  test.model <- DANA.test$data.model
  test.norm.model1 <- DANA.test$norm.models$test.TMM
  test.norm.model2 <- DANA.test$norm.models$test.RUVs

  # Set number of genes for negative correlation to minimum
  num.genes.plot.neg <- min(dim(test.model$neg$corr)[1],
                            dim(bench.model$neg$corr)[1],
                            sapply(DANA.test$norm.models, function(x) dim(x$neg$corr)[1]))

  # Subsetted correlation matrices
  corr.bench.neg <- bench.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.neg <- test.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm.neg <- lapply(DANA.test$norm.models, function(x) x$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg])

  ## Strictly positive correlations
  nnz.corr.bench.neg <- corr.bench.neg[upper.tri(corr.bench.neg)]
  nnz.corr.bench.neg <- nnz.corr.bench.neg[nnz.corr.bench.neg > 0]
  nnz.corr.bench.norm1.neg <- corr.bench.norm1.neg[upper.tri(corr.bench.norm1.neg)]
  nnz.corr.bench.norm1.neg <- nnz.corr.bench.norm1.neg[nnz.corr.bench.norm1.neg > 0]
  nnz.corr.test.neg <- corr.test.neg[upper.tri(corr.test.neg)]
  nnz.corr.test.neg <- nnz.corr.test.neg[nnz.corr.test.neg > 0]
  nnz.corr.test.norm.neg <- lapply(corr.test.norm.neg, function(x) x[upper.tri(x)][x[upper.tri(x)] > 0])
  names(nnz.corr.test.norm.neg) <- substr(names(nnz.corr.test.norm.neg), 6, 20)

>>>>

---
title: "DANA Normalization Assessment for the Paired MSK Data Sets"
author: "Yannick Dueren, Johannes Lederer, Li-Xuan Qin"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "results") })
---

@@

The settings list `settings` configure which parts, simulations, and analysis steps of this R markdown are executed. This way single components of the analysis can be disabled for the purpose of saving computation time.

```{r analysis.settings}
settings <- list(
  # Application of the full DANA pipeline to the benchmark data
  bench.DANA=T,
  # Application of the full DANA pipeline to the test data
  test.DANA=T,
  # Generate interactive plots of co-expression structures
  investigate.coexpression=F,
  # Comparison of coexpression estimation methods
  compare.corr.methods=T,
  # Full Pipeline comparison of two correlation methods
  compare.two.methods=F,
  # Compare DANA results for different cutoff choices
  compare.cutoffs=T,
  # Perform Differential Expression Analysis for test and benchmark data sets
  perform.DEA=T,
  # Generate paper figures
  generate.paper.figs=T,
  # Specify if paper figures are exported
  export.figures = T,
  # Path for file exports
  paper.fig.path = "../danawriteup/figs/",
  # Clears environment
  debug=TRUE
)
if(settings$generate.paper.figs) {
  settings$bench.DANA=TRUE
  settings$test.DANA=TRUE
}
if(settings$debug) rm(list=ls()[ls()!="settings"])
```

@@

@@

In this section, we compare two methods for estimating the co-expression of miRNAs to each other. The methods can be chosen individually to positive and negative controls and include all available marginal and partial correlation methods.

@@

```{r DANA.test.different.cutoffs}
if(settings$compare.cutoffs) {
  ## Configuration of three possible cutoff choices
  cutoff.conf1 <- list(
    t.zero = 2,  # zero-expressed in [0, t.zero)
    t.poor = 10,  # poorly-expressed in [t.zero, t.poor]
    t.well = 2^7  # well-expressed in [t.well, inf)
  )
  cutoff.conf2 <- list(
    t.zero = 1,  # zero-expressed in [0, t.zero)
    t.poor = 4,  # poorly-expressed in [t.zero, t.poor]
    t.well = 2^6  # well-expressed in [t.well, inf)
  )
  cutoff.conf3 <- list(
    t.zero = 3,  # zero-expressed in [0, t.zero)
    t.poor = 16,  # poorly-expressed in [t.zero, t.poor]
    t.well = 2^8  # well-expressed in [t.well, inf)
  )

@@

  # Combined read count histogram for test and benchmark data
  df <- data.frame(bench=bench.RC.log2, test=test.RC.log2)
  p.RC.hist <- ggplot(df) + 
    theme_classic() +
    geom_histogram(aes(x=test, y=..count..,fill="#69b3a2"), binwidth = .1) +
    geom_histogram(aes(x=bench, y=-..count.., fill="#404080"), binwidth = .1) +
    geom_vline(xintercept = log2(config$t.zero+1), color="#5851b8", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.poor+1), color="#5851b8", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.well+1), color="#E7298A",  linetype="longdash") +
    geom_segment(aes(x = log2(config$t.zero+1), y = 75, xend = log2(config$t.poor+1), yend = 75),
                 arrow=arrow(length=unit(.07, "in"), ends="both"),
                 color="#5851b8") +
    annotate(geom="label", x=(log2(config$t.zero+1)+log2(config$t.poor+1))/2.0, y=84,
             label="neg. contr.", color="#5851b8", size=4, fill = alpha(c("white"),0.85), label.size = NA) +
    geom_segment(aes(x = log2(config$t.well+1), y = 75, xend = 10, yend = 75),
                 arrow=arrow(length=unit(.07, "in"), ends="last"),
                 color="#E7298A") +
    annotate(geom="label", x=log2(config$t.well+1)+.5, y=84,
             label="pos. contr.", color="#E7298A", size=4, hjust=0, fill = alpha(c("white"),0.85), label.size = NA) +
    # ylim(c(-90,90)) +
    scale_fill_manual(name="Data set",values=c("#404080", "#69b3a2"), labels=c("benchmark", "test"), guide = guide_legend(reverse=TRUE)) +
    theme(legend.position=c(0.86,0.86), 
          # legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.9))) +
    scale_y_continuous(breaks=c(-50,0,50), labels=c(50,0,50), limits=c(-95,95)) +
    xlab("Mean log2(read count)") +
    ylab("Frequency")
  print(p.RC.hist)
  rm(df)

@@

  # Common control miRNAs
  pos.controls.bench <- rownames(bench.model$pos$corr)
  pos.controls.test <- rownames(test.model$pos$corr)
  pos.controls.common <- intersect(pos.controls.test, pos.controls.bench)
  neg.controls.bench <- rownames(bench.model$neg$corr)
  neg.controls.test <- rownames(test.model$neg$corr)

  # reduce the number of genes for the plot
  pos.controls.plot <- pos.controls.common[2:num.genes.plot.pos+1]

@@

  ### Generate plots
  # Positive controls
  p.corr.pos.bench <- ggplot.corr(corr.bench.pos, 
                                  clusters=clusters, 
                                  threshold=0, 
                                  title="benchmark (un-normalized)")
  p.corr.pos.test <- ggplot.corr(corr.test.pos, 
                                 clusters=clusters, 
                                 threshold=0, 
                                 title="test (un-normalized)")
  p.corr.pos.test.norm <- ggplot.corr(corr.test.norm.pos, 
                                     clusters=clusters, 
                                     threshold=0, 
                                     title="test (TMM-normalized)")

  p.corr.pos <- arrangeGrob(p.corr.pos.bench + theme(legend.position="none"), 
                            p.corr.pos.test + theme(legend.position="none"),
                            get.legend(p.corr.pos.bench), 
                            widths = c(2,2,1))
  grid.arrange(p.corr.pos)  # display plot

  p.corr.pos.three <- arrangeGrob(p.corr.pos.bench + theme(legend.position="none"), 
                                  p.corr.pos.test.norm + theme(legend.position="none"),
                                  p.corr.pos.test + theme(legend.position="none"),
                                  get.legend(p.corr.pos.bench + theme(legend.position = "bottom")),
                                  layout_matrix=rbind(c(1,2,3), c(4,4,4)),
                                  heights = c(5,1))
  grid.arrange(p.corr.pos.three)  # display plot

@@

  p.corr.neg <- arrangeGrob(p.corr.neg.bench + theme(legend.position="none"), 
                            p.corr.neg.test + theme(legend.position="none"),
                            get.legend(p.corr.neg.bench),
                            widths = c(2,2,1))
  grid.arrange(p.corr.neg)  # display plot

  p.corr.neg.three <- arrangeGrob(p.corr.neg.bench + theme(legend.position="none"), 
                                  p.corr.neg.test.norm + theme(legend.position="none"),
                                  p.corr.neg.test + theme(legend.position="none"),
                                  get.legend(p.corr.neg.bench + theme(legend.position = "bottom")),
                                  layout_matrix=rbind(c(1,1,2,2,3,3), c(4,4,4,4,4,4)),
                                  heights = c(5,1))
  grid.arrange(p.corr.neg.three)  # display plot

@@

```{r generate.paper.figs.corr.density.paper}
if(settings$generate.paper.figs) {
  # Co-expression models
  bench.model <- DANA.bench$data.model
  bench.norm.model1 <- DANA.bench$norm.models$benchmark.RUVr
  bench.norm.model2 <- DANA.bench$norm.models$benchmark.QN
  test.model <- DANA.test$data.model
  test.norm.model1 <- DANA.test$norm.models$test.RUVr
  test.norm.model2 <- DANA.test$norm.models$test.QN

  # Set number of genes for negative correlation to minimum
  num.genes.plot.neg <- min(dim(test.model$neg$corr)[1],
                            dim(bench.model$neg$corr)[1],
                            dim(test.norm.model1$neg$corr)[1],
                            dim(test.norm.model2$neg$corr)[1],
                            dim(bench.norm.model1$neg$corr)[1],
                            dim(bench.norm.model2$neg$corr)[1])

  # Subsetted correlation matrices
  corr.bench.neg <- bench.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.bench.norm1.neg <- bench.norm.model1$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.bench.norm2.neg <- bench.norm.model2$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.neg <- test.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm1.neg <- test.norm.model1$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm2.neg <- test.norm.model2$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]

  # Reduce to upper diagonal part
  corr.test.neg <- corr.test.neg[upper.tri(corr.test.neg)]
  corr.test.norm1.neg <- corr.test.norm1.neg[upper.tri(corr.test.norm1.neg)]
  corr.test.norm2.neg <- corr.test.norm2.neg[upper.tri(corr.test.norm2.neg)]
  corr.bench.neg <- corr.bench.neg[upper.tri(corr.bench.neg)]
  corr.bench.norm1.neg <- corr.bench.norm1.neg[upper.tri(corr.bench.norm1.neg)]
  corr.bench.norm2.neg <- corr.bench.norm2.neg[upper.tri(corr.bench.norm2.neg)]

  ## direct comparison of negative controls benchmark vs test vs test.norm1
  neg.corr <- data.frame(
    control=factor(c(rep("Benchmark", length(corr.bench.neg)), 
                     rep("Test", length(corr.test.neg)),
                     rep("Test (RUVr)", length(corr.test.norm1.neg)))),
    corr=c(corr.bench.neg, corr.test.neg, corr.test.norm1.neg)
    )

  p.density.neg.freq <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=.9) +
    theme_classic() +
    xlim(-1, 1) +
    scale_linetype_manual(values=c("solid", "solid", "twodash"))+
    scale_color_manual(values=c('#1B9E77','#D95F02','#D95F02')) + # First 2 colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.85,0.9), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation") +
    geom_vline(xintercept = 0, color="black", linetype="dashed")
  print(p.density.neg.freq)

  ## direct comparison of negative controls benchmark vs test vs test.norm1 vs test.norm2
  neg.corr <- data.frame(
    control=factor(c(rep("Benchmark", length(corr.bench.neg)), 
                     rep("Test", length(corr.test.neg)),
                     rep("Test (RUVr)", length(corr.test.norm1.neg)),
                     rep("Test (QN)", length(corr.test.norm2.neg)))),
    corr=c(corr.bench.neg, corr.test.neg, corr.test.norm1.neg, corr.test.norm2.neg)
    )

  p.density.neg.freq <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.075, alpha=.8, size=.9) +
    theme_classic() +
    xlim(-1, 1) +
    scale_color_manual(values=c('#e7298a','#1B9E77','#D95F02', '#7570B3')) + # First colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.85,0.86), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    theme(legend.key.width = unit(1.25,"cm")) +
    ylab("Frequency") +
    xlab("Marginal correlation")  +
    geom_vline(xintercept = 0, color="black", linetype="dashed")
  print(p.density.neg.freq)

  # Additional Plot with normalized benchmark
  neg.corr <- data.frame(
    control=factor(c(rep("Benchmark", length(corr.bench.neg)), 
                     rep("Benchmark (RUVr)", length(corr.bench.norm1.neg)),
                     rep("Benchmark (QN)", length(corr.bench.norm2.neg)),
                     rep("Test", length(corr.test.neg)),
                     rep("Test (RUVr)", length(corr.test.norm1.neg)),
                     rep("Test (QN)", length(corr.test.norm2.neg))),
                   levels=c("Benchmark", "Benchmark (RUVr)", "Benchmark (QN)", "Test", 
                            "Test (RUVr)", "Test (QN)")),
    corr=c(corr.bench.neg, corr.bench.norm1.neg, corr.bench.norm2.neg, corr.test.neg, corr.test.norm1.neg, corr.test.norm2.neg)
    )
  pallette <- brewer.pal(n=6, name="Paired")[c(1,2,4,5,6,8)]
  pallette[c(3,6)] <- c("cyan2", "magenta3")
  pallette <- c("steelblue3", "dodgerblue4", "cyan2", "indianred1", "red3", "magenta3")
  p.density.neg.freq.paper <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control, size=control)) + 
    geom_vline(xintercept = 0, size=.3, color="gray85", linetype="solid") +
    geom_freqpoly(binwidth=.05, alpha=1) +
    theme_classic() +
    xlim(-1, 1) +
    theme(legend.position=c(0.22,0.8), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5)),
          legend.key.width = unit(1.25,"cm")) +
    ylab("Frequency") +
    xlab("Marginal correlation") +
    scale_size_manual(values = c(.8,.4,.4,.8,.4,.4)) +
    scale_linetype_manual(values=c("solid", "twodash", "dashed", "solid", "twodash", "dashed")) +
    scale_color_manual(values=pallette)

  print(p.density.neg.freq.paper)

  if(settings$export.figures) {
    # ggsave(paste0(settings$paper.fig.path, "MSK_CorrDensityNeg_Hist.pdf"), p.density.neg.hist, width = 5, height=4, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "MSK_CorrDensityNeg_Freq.pdf"), p.density.neg.freq, width = 5, height=4, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "MSK_CorrDensityNeg_Freq_WithBench.pdf"), p.density.neg.freq.paper, width = 5, height=4, device="pdf")
  }
}
```

@@

```{r generate.paper.figs.corr.density.supplements.test}
if(settings$generate.paper.figs) {
  models <- list(
    "test" = DANA.test$data.model,
    "benchmark" = DANA.bench$data.model,
    "test (TMM)" = DANA.test$norm.models$test.TMM,
    "test (DESeq)" = DANA.test$norm.models$test.DESeq,
    "test (PoissonSeq)" = DANA.test$norm.models$test.PoissonSeq,
    "test (RUVg)" = DANA.test$norm.models$test.RUVg,
    "test (RUVs)" = DANA.test$norm.models$test.RUVs,
    "test (RUVr)" = DANA.test$norm.models$test.RUVr,
    "test (TC)" = DANA.test$norm.models$test.TC,
    "test (Med)" = DANA.test$norm.models$test.Med,
    "test (UQ)" = DANA.test$norm.models$test.UQ,
    "test (QN)" = DANA.test$norm.models$test.QN
  )

  # Set number of genes for negative correlation to minimum found
  num.genes.plot <- min(sapply(models, function(x) dim(x$neg$corr)[1]))

  # Subsetted correlation matrices
  corrs <- lapply(models, function(x) x$neg$corr[1:num.genes.plot, 1:num.genes.plot])
  corrs <- lapply(corrs, function(x) x[upper.tri(x)])

  control <- factor(
    as.vector(sapply(names(corrs), function(x) rep(x,length(corrs[[1]])))),
    levels = c("test","benchmark","test (TMM)","test (DESeq)","test (PoissonSeq)","test (RUVg)","test (RUVs)","test (RUVr)","test (TC)","test (Med)","test (UQ)", "test (QN)")
  )
  neg.corr <- data.frame(
    control=factor(control),
    corr=unname(unlist(corrs))
  )
